Model Finding for Recursive Functions in SMT

### 2016-04-12 ijcar2016/body.tex
Introduced restriction to "Herbrand" theories

A \emph{$\Sigma$-interpretation $\I$} %is a mathematical structure that
maps each type $\tau \in \stypes{\Sigma}$ to a nonempty set $\tau^\I$,
the \emph{domain} of~$\tau$ in~$\I,$
each function symbol $\con{f} : \tau_1 \times \cdots \times \tau_n \rightarrow \tau$ in $\sfuns{\Sigma}$
to a total function $\con{f}^\I : \tau_{\smash{1}}^\I \times \cdots \times\allowbreak \tau_n^\I \rightarrow\allowbreak \tau^\I$,
and each variable $x:\tau$ of $\svars{\Sigma}$ to an element of $\tau^\I$.
A \emph{theory} is a pair $T = (\Sigma, \Mo)$ where
$\Sigma$ is a signature and $\Mo$ is a class of $\Sigma$-interpretations,
the \emph{models} of $T$, closed under variable reassignment
(i.e., for every $\I \in \Mo$, every $\Sigma$-interpretation that differs
from $\I$ only on the variables of $\svars \Sigma$ is also in $\Mo$).
A $\Sigma$-formula $\varphi$ is \emph{$T$-satisfiable}
if it is satisfied by some interpretation in $\Mo$,
which we call a \emph{$T$-model of $\varphi$}\begin{rep},
otherwise, it is \emph{$T$-unsatisfiable}\end{rep}.
A formula $\varphi$ \emph{$T$-entails} $\psi$, written $\varphi \models_T \psi$,
if all interpretations in $\Mo$ that satisfy $\varphi$ also satisfy $\psi$.
Two formulas $\varphi$ and $\psi$ are \emph{$T$-equivalent}
if each $T$-entails the other.
If $T_0 = (\Sigma_0, \Mo_0)$ is a theory and $\Sigma$ is a signature
with $\sfuns{\Sigma_0} \subseteq \sfuns{\Sigma}$ 
and $\stypes{\Sigma_0} \subseteq \stypes{\Sigma}$,
the \emph{extension of $T_0$ to $\Sigma$} is the theory $T = (\Sigma, \Mo)$ 
where $\Mo$ is the set of all $\Sigma$-interpretations $\I$
whose $\Sigma_0$-reduct is a model of $T_0$.
Note that $T$ is a \emph{conservative} extension of $T_0$, in the sense that
a $\Sigma_0$-formula is $T_0$-satisfiable if and only it is $T$-satisfiable.
We refer to the symbols of $\Sigma$ that are not in $\Sigma_0$ as \emph{uninterpreted}.

For the rest of the paper, \relax{we fix a theory $T = (\Sigma, \Mo)$
with uninterpreted symbols}, constructed as above.

>>>>

A \emph{$\Sigma$-interpretation $\I$} %is a mathematical structure that
maps each type $\tau \in \stypes{\Sigma}$ to a nonempty set $\tau^\I$,
the \emph{domain} of~$\tau$ in~$\I,$
each function symbol $\con{f} : \tau_1 \times \cdots \times \tau_n \rightarrow \tau$ in $\sfuns{\Sigma}$
to a total function $\con{f}^\I : \tau_{\smash{1}}^\I \times \cdots \times\allowbreak \tau_n^\I \rightarrow\allowbreak \tau^\I$,
and each variable $x:\tau$ of $\svars{\Sigma}$ to an element of $\tau^\I$.
A \emph{theory} is a pair $T = (\Sigma, \Mo)$ where
$\Sigma$ is a signature and $\Mo$ is a class of $\Sigma$-interpretations,
the \emph{models} of $T$, closed under variable reassignment
(i.e., for every $\I \in \Mo$, every $\Sigma$-interpretation that differs
from $\I$ only on the variables of $\svars \Sigma$ is also in $\Mo$).
A $\Sigma$-formula $\varphi$ is \emph{$T$-satisfiable}
if it is satisfied by some interpretation in $\Mo$,
which we call a \emph{$T$-model of $\varphi$}\begin{rep},
otherwise, it is \emph{$T$-unsatisfiable}\end{rep}.
A formula $\varphi$ \emph{$T$-entails} $\psi$, written $\varphi \models_T \psi$,
if all interpretations in $\Mo$ that satisfy $\varphi$ also satisfy $\psi$.
Two formulas $\varphi$ and $\psi$ are \emph{$T$-equivalent}
if each $T$-entails the other.
We call a theory a \emph{Herbrand} theory if 
for every quantifier-free $\Sigma$-formula $\psi$ over the variables $\vec x$,
$\{\psi[\vec t/\vec x] \mid t \text{ ground $\Sigma$-term}\} \models_T \forall \vec x.\; \psi$.

If $T_0 = (\Sigma_0, \Mo_0)$ is a theory and $\Sigma$ is a signature
with $\sfuns{\Sigma_0} \subseteq \sfuns{\Sigma}$ 
and $\stypes{\Sigma_0} \subseteq \stypes{\Sigma}$,
the \emph{extension of $T_0$ to $\Sigma$} is the theory $T = (\Sigma, \Mo)$ 
where $\Mo$ is the set of all $\Sigma$-interpretations $\I$
whose $\Sigma_0$-reduct is a model of $T_0$.
Note that $T$ is a \emph{conservative} extension of $T_0$, in the sense that
a $\Sigma_0$-formula is $T_0$-satisfiable if and only it is $T$-satisfiable.
We refer to the symbols of $\Sigma$ that are not in $\Sigma_0$ as \emph{uninterpreted}.

For the rest of the paper, \relax{we fix a theory $T = (\Sigma, \Mo)$
with uninterpreted symbols}, constructed as above, and assume it is a Herbrand theory.
While this is an actual restriction, it can be shown that the theories typically considered 
in SMT are Herbrand.
